Ancestry linkage in hybrid zones

### 2015-12-09 drafts/hybrid_zone.tex
added to draft

\section{Introduction}
The process of speciation commonly involves allopatric divergence in relative isolation \cite{Coyne2004}. When formerly isolated populations come back into contact, a gradient of ancestry can form in geographic space, resulting in a hybrid zone \cite{Barton1985}. Throughout the genome, clines in ancestry may diminish over time due to migration; however, subsets of the genome that are experiencing differential selection \alisa{better way to say this?} may be expected to maintain clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \cite{Barton1979a}. The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which diminishes with physical genomic distance \cite{Barton1986}.\\

Many population genetic models of hybrid zones have focused on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}. These models predict cline width to be narrowest at and nearby selected loci, with cline width at unliked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow. Population genetic patterns at these loci are important because, in practice, determining whether a hybrid zone has reached equilibrium can be challenging. In particular, even if clines at selected loci are stable, the rest of the genome may still contain additional information.

\alisa{Are we mainly just arguing that we get more robust information from looking at linkage around selected loci? Is there anything else?}

In addition to patterns of allele frequency, the extent of linkage in hybrid zones can also provide information about important parameters, such as age and rates of gene flow, in hybrid zones. As denser genotype data becomes more readily available, it is now possible to utilize patterns of linkage in the analysis of hybrid zones. Genome-wide covariance in ancestry can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore can provide information about the age of the hybrid zone, and the strength of selection at nearby loci.\\

Using an equilibrium model. We often don't know if hybrid zones are at equilibrium.

@@

%-Studying hybrid zones is a great opportunity to understand the maintenance of distinct species, and the process of speciation. Here we consider hybrid zones that are a cline that is maintained by selection against hybrids (so-called tension zone). \\
%-Gene flow across a hybrid zone may be impeded by selection against migrants. In older contact zones, loci that are associated with reduced fitness in heterozygotes will have steeper clines than the rest of the genome. There is probably a lot to learn from geographic patterns at loci neighboring selected loci (e.g. small block lengths indicate that there has been plenty of opportunity for recombination in heterozygous backgrounds)\\ 
%-There has been a lot of investigation into the equilibrium state of hybrid zones, at both selected and linked loci. \\
-We consider the non-equilibrium state of hybrid zones, and patterns of linkage around selected loci. This will become increasingly relevant as studiers of HZs get genomic data and can reliably estimate block lengths. \\

@@

\subsection{Data}
We compare the predictions of our model to genomic data obtained from individuals sampled a naturally occurring hybrid zone between \emph{Mus m. musculus/M. m. domesticus} \cite{Turner2011,Turner2014}. The genomic data generated by \citep{Turner2014} was for a mapping population generated by mating wild-caught individuals. Because some individuals were mated multiple times, the mapping population comprises many closely related individuals. We approximate wild population samples by restricting our analysis to genotyped individuals that had both parents sampled in the same location in such a way that none of the individuals shared a parent (see table for list of individuals).\\

Ancestry blocks were estimated using ChromopainterV2 \cite{Lawson2012}.

\end{document}

>>>>

\section{Introduction}
The process of speciation commonly involves allopatric divergence in relative isolation \cite{Coyne2004}. When formerly isolated populations come back into contact, a gradient of ancestry can form in geographic space, resulting in a hybrid zone \cite{Barton1985}. Throughout the genome, clines in ancestry may diminish over time due to migration; however, subsets of the genome that are experiencing selection may be expected to maintain clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \cite{Barton1979a}. The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which diminishes with physical genomic distance \cite{Barton1986}.\\

Many population genetic models of hybrid zones have focused on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}. These models predict cline width to be narrowest at and nearby selected loci, with cline-width at unliked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow. As denser genotype data becomes more readily available, we can begin to utilize patterns of linkage in the analysis of hybrid zones. Genome-wide covariance in ancestry can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone, and the strength of selection at linked loci. It is our goal here to understand how strength of selection, density of selected sites, recombination rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations. This may be especially important for an understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctions. In the extreme case of zero fitness in first-generation hybrids, for example, paternal genomes will never mix and the stable hybrid zone will resemble a recently formed one. If hybrids have non-zero fitness, and some amount of introgression can occur, and patterns of introgression can provide information about the strength of selection, extent of hybridization, and timing of secondary contact. In this latter case, it may be of additional interest to establish whether the hybrid zone is in its equilibrium state. \\

@@

%-Studying hybrid zones is a great opportunity to understand the maintenance of distinct species, and the process of speciation. Here we consider hybrid zones that are a cline that is maintained by selection against hybrids (so-called tension zone). \\
%-Gene flow across a hybrid zone may be impeded by selection against migrants. In older contact zones, loci that are associated with reduced fitness in heterozygotes will have steeper clines than the rest of the genome. There is probably a lot to learn from geographic patterns at loci neighboring selected loci (e.g. small block lengths indicate that there has been plenty of opportunity for recombination in heterozygous backgrounds)\\ 
%-There has been a lot of investigation into the equilibrium state of hybrid zones, at both selected and linked loci. \\

@@

\subsection{Data}
We compare the predictions of our model to genomic data obtained from individuals sampled a naturally occurring hybrid zone between \emph{Mus m. musculus/M. m. domesticus} \cite{Turner2011,Turner2014}. The genomic data generated by \citep{Turner2014} was for a mapping population generated by mating wild-caught individuals. Because some individuals were mated multiple times, the mapping population comprises many closely related individuals. We approximate wild population samples by restricting our analysis to genotyped individuals that had both parents sampled in the same location in such a way that none of the individuals shared a parent (see table for list of individuals). Approximately 280,000 SNPs were represented in the final dataset.

Ancestry blocks were estimated using ChromopainterV2 \cite{Lawson2012}. \\

Although we expect the age of this hybrid zone to correspond to historical human movement, and therefore be thousands of years old \cite{Teschke2008}, weighted the physical scale of ancestry-disequilibrium (both in the form of weighted ancestry-LD \citep{Loh2013} and ancestral block lengths) suggest relatively recent hybridization. Specifically, the high LD between quite distantly linked SNPs in the FS population suggests that a relatively high proportion of early-generation hybrids, whereas the rapid decay of LD in the the TU population reflects a much older hybridization event. Importance of geographic distribution of popgen patterns.

\begin{table}
\begin{tabular}{l|c|c}
Individual ID \citep{Turner2014} & Population \citep{Turner2011} & Longitude(E)\\
\hline
FP112 & FS&11.665\\
FP67 & FS &11.665\\
FP114 & FS &11.665\\
FP79 & FS &11.665\\
FP152 &  FS &11.665\\
FP11 & GL &11.965\\
FP111&GL &11.965\\
FP59 & HA  &11.721\\
FP92 & HA &11.721\\
FP7 &  HA  & 11.721\\
FP45 & HA&11.721\\
FP244& HO &11.693\\
FP9&   RE&11.994\\
FP5 &  RF&11.767\\
FP78 &  SO &11.539\\
FP52 &  SO &11.539\\
FP180& SO &11.539\\
FP14 & ST&11.539\\
FP40 & ST&11.539\\
FP145 & TU&11.748\\
FP29&  TU&11.748\\
FP135& TU&11.748\\

\end{tabular}
\end{table}
\end{document}
